---
title: "Project_Paper_Group6"
output: html_document
date: "2024-12-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include = FALSE}
library(ISLR)
library(tidyverse) 
library(tidymodels)
library(knitr)
library(rpart.plot)
library(caret)
library(dplyr)
library(dials)
```

# Drum Sample Machine Learning Classifier

## Motivation


The motivation behind this project was our love for music and working with audio. For me personally, I produce a lot of music and have my own library of drum sounds that I pulled for this project. I would like to eventually use machine learning techniques on my own music to streamline the production and song-writing process. I have a vision of streaming a track to an AI Agent / database, which segments each part of the song (Vocals, Guitars, Bass, Drums, etc.) and gives suggestions based on the context of the song/database.

Creating a drum classification tool seemed like a good first step to creating an app like this, teaching skills like working with audio data and data manipulation. Working with audio data is not very well doccumented online as it is a niche interest, so this project provides the opportunity to learn skills that interest me!

## Data Preperation

### Data Collection

The wav files were pulled from a personal library of drum sounds, primarily from Splice. They were automatically filtered based on regex patterns in the file name. The features for each sample were extracted using the CREPE library (dicussed in the next section) manually using local hardware, and were sabed into a CSV titled 'sample_features.csv'. As we found out later, some of the data was messy due to the regex method of creating the dataset, so some data filtering was necessary.

{Plot Distribution of samples}

### Data Cleaning

To increase accuracy and omit any outliers, we filtered samples that were longer than 10 seconds (implying the sample may be a loop instead of a one-shot). And further filtered kicks, hats, and snares to be less than 5 seconds. The ride and crash labels were combined into one catecory, We also removed some excess labels, such as perussion, claps, and toms. Out initial generated dataset was very unballanced, shown below.

Initial dataset

kick     471
snare    467
tom      250
clap     138
ride     120
hat      111


Because the accuracies for the different labels varied greatly instrument to instrument, we decided to balance the dataset based on the minimum number of samples in a given category, which was rides. After balancing, the dataset is much smaller, but is also less biased, so we decided to use it for this project.

After filtering + balancing:

hat      103
kick     103
ride     103
snare    103


### Feature Engineering

In order to extract features from the dataset, we used a python library called 'crepe', which is a monophonic pitch tracker based on a deep convolutional neural network that operates directly on the time-domain waveform input {https://pypi.org/project/crepe/}.

Features: * Core Frequency: Finds frequency thorughout the time-series-sample that has the highest overall confidence {Core Freq Plot} * Max Amplitude: Finds the maximum amplitude (loudness) {Max Amp Plot} * Average Activation: ??? {Activation Plot} * Duration: Length of the sample {Duration Plot}

## Methodologies

### Tree Model

```{r}
drums <- read.csv("sample_features_balanced.csv") |>
  drop_na() |>
  mutate(label = as.factor(label)) 

```



#### Initial Tree Model

The drums data set is split into 75% training data, and 25% testing data, stratified by the classification 'label'. A basic tree model is fit using the decision_tree() function with five terminal nodes.  

The tree has four decisions: 
- core frequency < 111
- duration < 1.5
- average activation < 0.0052
- max amplitude < 0.57

```{r}
set.seed(445)

data_split <- initial_split(drums, 0.75, strata = label)
training_drums <- training(data_split)
testing_drums <- testing(data_split)

tree_spec <- decision_tree() |>
  set_engine('rpart') |>
  set_mode("classification")

tree_fit <- tree_spec |>
  fit(label ~. , data = training_drums)

tree_engine <- extract_fit_engine(tree_fit)
rpart.plot(tree_engine, roundint =  FALSE)
```

#### Predictions and Accuracy

The initial model has a testing accuracy of 0.86 and correctly classifies:

- 22/26 Hats
- 25/26 Kicks
- 25/26 Rides
- 18/26 Snares

```{r, echo = FALSE}
set.seed(445)

predictions <- tree_fit |>
  predict(testing_drums) |>
  pull(.pred_class)

metrics <- metric_set(accuracy)

model_performance <- testing_drums %>%
 mutate(predictions = predictions) %>%
 metrics(truth = label, estimate = predictions)

confmatrix <- confusionMatrix(testing_drums$label,predictions)
confmatrix$table


model_performance$.estimate
```

#### Tuning Initial Tree Model

We used 10-fold cross-validation to find the best cost_complexity parameter and tree depth. 

- Best CC Paremeter: 0.01
- Best tree depth: 5

```{r, echo = FALSE}
set.seed(445)

tune_spec <- 
  decision_tree(
    cost_complexity = tune(),
    tree_depth = tune()
  ) %>% 
  set_engine("rpart") %>% 
  set_mode("classification")


tree_grid <- grid_regular(cost_complexity(),
                          tree_depth(),
                          levels = 10)


folds <- vfold_cv(training_drums, v = 10)

tree_wf <- workflow() %>%
  add_model(tune_spec) %>%
  add_formula(label ~ .)

tree_res <- tree_wf %>% 
  tune_grid(resamples = folds,
            grid = tree_grid)



tree_res %>%
  collect_metrics() %>%
  dplyr::filter(.metric != "brier_class") %>%
  dplyr::filter(.metric != "roc_auc") %>%
  mutate(tree_depth = factor(tree_depth)) %>%
  ggplot(aes(cost_complexity, mean, color = tree_depth)) +
  geom_line(linewidth= 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) +
  scale_x_log10(labels = scales::label_number()) +
  scale_color_viridis_d(option = "plasma", begin = .9, end = 0)



best_tree <- tree_res %>%
  select_best(metric="accuracy")

tuned_cc <- best_tree$cost_complexity
tuned_cc
best_tree$tree_depth
```

#### Final Tree Model

We refit the tree model using our tuned cost complexity parameter. With the new tuned paramter, we produce the exact same model and predictions as oour initial model. 

```{r, echo = FALSE}
set.seed(445)

tree_spec_tuned <- decision_tree(cost_complexity = tuned_cc) |>
  set_engine('rpart') |>
  set_mode("classification")

tree_fit_tuned <- tree_spec_tuned |>
  fit(label ~. , data = training_drums)


predictions_tuned <- tree_fit_tuned |>
  predict(testing_drums) |>
  pull(.pred_class)

metrics_tuned <- metric_set(accuracy)

model_performance_tuned <- testing_drums %>%
 mutate(predictions = predictions_tuned) %>%
 metrics_tuned(truth = label, estimate = predictions)


model_performance_tuned$.estimate

confmatrix <- confusionMatrix(testing_drums$label,predictions_tuned)
confmatrix$table

tree_engine_tuned <- extract_fit_engine(tree_fit_tuned)
rpart.plot(tree_engine_tuned, roundint =  FALSE)

```

#### Discussion

Although our model did not change at all after tuning the cost complexity parameter, we can explain this with our graph of the 10-fold cross validation step. We can see that changing the tree depth does almost nothing to change our accuracy for all trees with depth higher than 2. The cost complexity parameter has no effect on the accuracy until it has a value of 0.01 or higher, where there is a slight increase in accuracy. 


### Olivia 


### Aaron


## Issues
The first issue we ran into was feature selection. The issues we ran into were mainly low accuracies from the original dataset.

## Improvements
Through data cleaning techniques, we were able to greatly improve the accuracy of the model. Initially, we were getting accuracies of around 60%, which we were able to improve to 88% with a Random Forrest Model.

{Include Tables comparing the initial results before the dataset was cleaned}

Model after improvements
{Include Tables comparing the initial results after the dataset was cleaned}

Further Analysis
PCA Visualization of feature space to explain support vector

Conclusion
References
