---
title: "Introduction"
output: html_document
---

Load data 

```{r}
library(readr)
library(ggplot2)
library(dplyr)
library(tidyr)


bank_full <- read_delim("data/bank-full_1.csv", delim = ';', escape_double = FALSE, trim_ws = TRUE)
head(bank_full)
```

The data set consists of 16 variables describing the customer and marketing campaign along with temporal aspects:
* age: Age in years 
* job: Occupation (qualitative variable)
* marital: Marital status ( qualitative variable)
* education: Highest level of education (qualitative variable)
* default: has credit in default? (binary)
* balance: average yearly balance
* housing: has housing loan? (binary)
* loan: has personal loan? (binary)
* Contract: contact communication type (qualitative)
* day: last contact day of the week (month)
* month: last contact month of year
* duration: last contact duration, in seconds
* campaign: number of contacts performed during this campaign and for this client
* pdays: number of days that passed by after the client was last contacted from a previous campaign (-1 means client was not previously contacted)
* previous: number of contacts performed before this campaign and for this client
* poutcome: outcome of the previous marketing campaign (qualitative)
* y: Indicator of whether the client subscribed a term deposit (Binary target) 



## Distribution

```{r}
library(tidyverse)
library(ggplot2)

bank_long <- bank_full %>%
  select(where(is.numeric)) %>% # numeric columns
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value")

ggplot(bank_long, aes(x = value)) +
  geom_histogram(bins = 30, fill = "darkblue", color = "white", alpha = 0.7) +
  facet_wrap(~ variable, scales = "free", ncol = 4) + 
  labs(
    title = "Distribution of Numeric Variables in bank_full Dataset",
    x = "Value",
    y = "Frequency"
  ) +
  theme_minimal()

```


From the plots one can observe that the numeric variables are heavily skewed (balance, campaign, duration, pdays, previous). The dissemblance of the observations should take into account when considering the bias in our model fit.

```{r}

# Calculate and print the success rate
success_rate <- mean(bank_full$y == "yes") * 100
cat("Overall Subscription Success Rate:", round(success_rate, 2), "%\n")

# bar plot for the target variable 'y'
ggplot(bank_full, aes(x = y, fill = y)) +
  geom_bar() +
  labs(title = "Distribution of Term Deposit Subscription (y)",
       x = "Subscribed to Term Deposit?",
       y = "Count") +
  theme_minimal()

```

**Description:** A simple bar chart showing the imbalance between clients who subscribed to a term deposit (“yes”) and those who did not (“no”), with the majority belonging to the “no” category.



# Convert variables to factors for plotting

```{r}

bank_full$y <- as.factor(bank_full$y)
bank_full$job <- as.factor(bank_full$job)
bank_full$marital <- as.factor(bank_full$marital)
bank_full$education <- as.factor(bank_full$education)
bank_full$contact <- as.factor(bank_full$contact)
bank_full$month <- as.factor(bank_full$month)
bank_full$poutcome <- as.factor(bank_full$poutcome)
bank_full$housing <- as.factor(bank_full$housing)
bank_full$loan <- as.factor(bank_full$loan)

```


```{r}

ggplot(bank_full, aes(x = job, fill = y)) +
  geom_bar(position = "fill") +
  coord_flip() +
  labs(title = "Subscription Rate by Job",
       y = "Proportion") +
  theme_minimal()

```

**Description:** A stacked bar chart illustrating the proportion of subscription outcomes across different job types, with retired and student occupations showing relatively higher subscription rates.


```{r}
ggplot(bank_full, aes(x = education, fill = y)) +
  geom_bar(position = "fill") +
  coord_flip() +
  labs(title = "Subscription Rate by Education Level",
       y = "Proportion") +
  theme_minimal()
```

**Description:** A proportional bar chart comparing subscription rates across education levels, indicating that clients with tertiary education have a higher tendency to subscribe.


```{r}

ggplot(bank_full, aes(x = marital, fill = y)) +
  geom_bar(position = "fill") +
  coord_flip() +
  labs(title = "Subscription Rate by Marital Status",
       y = "Proportion") +
  theme_minimal()

```


**Description:** A stacked bar chart showing subscription proportions by marital status, revealing that single clients have a slightly higher subscription rate compared to married or divorced clients.


```{r}
ggplot(bank_full, aes(x = contact, fill = y)) +
  geom_bar(position = "fill") +
  labs(title = "Subscription Rate by Contact Type",
       y = "Proportion") +
  theme_minimal()
```


**Description:** A proportional bar chart comparing subscription outcomes by contact method, with cellular contacts showing a notably higher subscription rate than telephone or unknown contact types.


```{r}
ggplot(bank_full, aes(x = poutcome, fill = y)) +
  geom_bar(position = "fill") +
  labs(title = "Subscription Rate by Previous Marketing Campaign Outcome",
       y = "Proportion") +
  theme_minimal()
```

**Description:** A stacked bar chart depicting how the outcome of previous marketing campaigns influences current subscription rates, with previous “success” outcomes leading to higher subscription proportions.



```{r}
bank_full %>%
  group_by(campaign) %>%
  summarise(success_rate = mean(y == "yes")) %>%
  ggplot(aes(x = campaign, y = success_rate)) +
  geom_line() +
  geom_point() +
  labs(title = "Subscription Success Rate vs. Number of Campaign Contacts",
       x = "Number of Contacts During This Campaign",
       y = "Success Rate") +
  theme_minimal()
```

**Description:** A line plot with points showing how the subscription success rate declines as the number of contacts during the campaign increases, suggesting diminishing returns with repeated contact attempts.


```{r}
summary(bank_full)
```

The categorical variables need to be set to categorical (or binary) variables, since R is reading the data as mere strings.

```{r}
bank_full$job <- as.factor(bank_full$job)
bank_full$marital <- as.factor(bank_full$marital)
bank_full$education <- as.factor(bank_full$education)
bank_full$contact <- as.factor(bank_full$contact)
bank_full$poutcome <- as.factor(bank_full$poutcome)
bank_full$day <- as.factor(bank_full$day)
bank_full$month <- as.factor(bank_full$month)

bank_full$default <- ifelse(bank_full$default == "yes", 1, 0)
bank_full$housing <- ifelse(bank_full$housing == "yes", 1, 0)
bank_full$loan <- ifelse(bank_full$loan == "yes", 1, 0)

```


Plots for categorical variables (S.A.)

```{r}

cat_vars <- c("job", "marital", "education", "contact", "month", "poutcome")

for (var in cat_vars) {
p <- ggplot(bank_full, aes_string(x = var, fill = "y")) +
geom_bar(position = "fill") +
coord_flip() +
labs(
title = paste("Proportion of Subscription by", var),
x = var,
y = "Proportion"
) +
theme_minimal()

print(p) 
}

```
**Description:** These plots show that subscription rates vary meaningfully across categorical variables, with certain jobs, higher education levels, specific contact methods (especially cellular), particular months, and prior campaign successes all associated with higher proportions of clients subscribing to a term deposit.
