---
title: "model_fit"
output: html_document
---

load libraries and clean data

```{r}
library(readr)
library(tidyverse)
library(boot)
library(glmnet)
library(caret)
library(pROC)
library(ggplot2)

set.seed(445) 

bank_full <- read_delim("data/bank-full_1.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)
```


```{r}
bank_full$y <- factor(bank_full$y, levels = c("no", "yes"))


bank_full <- bank_full %>%
  mutate(
    job = as.factor(job),
    marital = as.factor(marital),
    education = as.factor(education),
    default = as.factor(default),
    housing = as.factor(housing),
    loan = as.factor(loan),
    contact = as.factor(contact),
    month = as.factor(month),
    poutcome = as.factor(poutcome))

```


## Train/Test split
- Explain purpose of an honest out-of-sample evaluation.
- k-fold CV, k=5

```{r}
bank_full$previously_contacted <- factor(ifelse(bank_full$pdays == -1, "No", "Yes"), levels = c("No", "Yes"))
bank_full$pdays_non_miss <- ifelse(bank_full$pdays == -1, NA, bank_full$pdays)

train_index <- createDataPartition(bank_full$y, p = 0.8, list = FALSE, times = 1)
train_data <- bank_full[train_index, ]
test_data <- bank_full[-train_index, ]
```


## subset selection
- Backward step-wise elimination (ecalutes fit using AIC)

```{r}
initial_formula <- y ~ age + job + marital + education + default + balance + housing + loan + 
  contact + day + month + campaign + pdays + previous + poutcome + previously_contacted

# Fit the model with maximum covaraints on the training set
glm.full <- glm(initial_formula, data = train_data, family = binomial)

# Use the step function to perform backward elimination based on AIC
glm.selected <- step(
  glm.full, 
  direction = "backward", 
  trace = 0
)

selected_formula <- formula(glm.selected)
selected_formula

```


    

```{r}

interact_formula <- update(selected_formula, . ~ .^2) 

# Fit the maximal interaction model (CAUTION: This can be slow and unstable with many factors)
glm.full.interact <- glm(interact_formula, data = train_data, family = binomial)
glm.full.interact.summary <- summary(glm.full.interact)
```


```{r}
coef_table <- glm.full.interact.summary$coefficients
coef_df <- as.data.frame(coef_table)

view(coef_df %>%  filter(`Pr(>|z|)`< 0.01))
```

job:contact
job:month
education:housing
housing:month
contact:month
day:month
month:poutcome


```{r}
interact_formula_significant <- y ~ job + marital + education + balance + housing + loan + contact + 
  day + month + campaign + previous + poutcome + job:contact + job:month + education:housing + housing:month +
  contact:month + day:month + month:poutcome
  
glm.full.interact_significant <- glm(interact_formula_significant, data = train_data, family = binomial)


glm.interact.selected <- step(
  glm.full.interact_significant, 
  direction = "backward", 
  trace = 1 
)

selected_interact_formula <- formula(glm.interact.selected)
selected_interact_formula
```



---

# 3. Modeling 

## 3.1 Logistic Regression (Baseline Model)
- Fit logistic regression.
- k-fold CV for estimated test error.
- Assess linearity and consider interactions if appropriate.
- Optional: **regularized logistic regression** (Ridge/Lasso) using CV-selected $\lambda$.



*Standard raw logistic reg (without tuning)*

```{r}
glm_out <- glm(y ~ ., data = bank_full, family = binomial)
summary(glm_out)
```

Discuss accuracy (cv test error) and amount of parameters

```{r}
glm.full.cv <- glm(initial_formula, data = data, family = binomial, x = TRUE)

cv.error.full <- cv.glm(data, glm.full.cv, K = 5)$delta[2]
length(coef(glm.full.cv))
```

This model is complex due to the large number of dummy variables. The CV test error provides an estimate of how well this complexity generalizes to new, out of sample data. If the CV error is significantly higher than the training error, the model is likely overfitting


*With step-wise feature selection* (our base line)

```{r}
# Re-fit the final selected model
glm.baseline <- glm(selected_formula, data = train_data, family = binomial)
summary(glm.baseline)

```


```{r}
glm_out <- glm(y ~ job + marital + education + balance + housing + loan + contact + 
    day + month + campaign + previous + poutcome,
    data = bank_full,
    family = binomial, # Or whatever family you used
    x = TRUE,
    model = TRUE
)

cv.error.baseline <- cv.glm(
    bank_full, #only include features that are fit
    glm_out, K = 5)

cv.error.baseline$delta[2]
```


```{r}
glm.probs.baseline <- predict(glm.baseline, newdata = test_data, type = "response")

glm.pred.baseline <- factor(ifelse(glm.probs.baseline > 0.5, "yes", "no"), levels = c("no", "yes"))
confusionMatrix(glm.pred.baseline, test_data$y, positive = "yes")

```


- analysis of parameters and example predictions with interpretation


residual plots (numeric features)... this will lead to inclusion of non-linear features

```{r}
glm.resids <- residuals(glm.full, type = "deviance")

# Function to plot residuals vs a numeric predictor
plot_residuals_vs_predictor <- function(predictor_name, data, residuals) {
  df <- data.frame(
    Residuals = residuals,
    Predictor = data[[predictor_name]]
  ) %>% filter(!is.na(Predictor))
  
  p <- ggplot(df, aes(x = Predictor, y = Residuals)) +
    geom_point(alpha = 0.3) +
    geom_smooth(method = "loess", color = "red", se = FALSE) + # Add a smoothing line
    geom_hline(yintercept = 0, linetype = "dashed", color = "blue") +
    labs(
      title = paste("Deviance Residuals vs.", predictor_name),
      y = "Deviance Residuals",
      x = predictor_name
    ) +
    theme_minimal()
  print(p)
}

# Plot residuals for key numeric features
plot_residuals_vs_predictor("age", train_data, residuals(glm.full, type = "deviance"))
plot_residuals_vs_predictor("balance", train_data, residuals(glm.full, type = "deviance"))
```





*polynomials logit reg (on Balance or other numeric variables)*
- test models with transformations (polynomial or Log)
- test models with interaction terms

```{r}

```


```{r}
glm.test.linearity <- glm(y ~ job + marital + education + balance + housing + loan + contact + 
    day + month + campaign + previous + poutcome + I(balance^2), data = train_data, family = binomial)

summary(glm.test.linearity)$coefficients["I(balance^2)", "Pr(>|z|)"]

# Interaction: Test interaction between 'age' and 'campaign' (common drivers).
glm.test.interaction <- glm(selected_formula, data = train_data, family = binomial)
glm.test.interaction <- update(glm.test.interaction, . ~ . + age:campaign)

print(anova(glm.full, glm.test.interaction, test = "Chisq"))
```

*Regularized logit reg*

Ridge
- test across lambda values (and plot) with mean CV test error
- confussion matrix and roc curve

```{r}
full_predictors_formula <- y ~ age + job + marital + education + default + balance + housing + loan + 
  contact + day + month + campaign + pdays + previous + poutcome + previously_contacted

X_train_ridge <- model.matrix(full_predictors_formula, data = train_data)[, -1] 
y_train_ridge <- train_data$y

```

k-fold CV for optimal lambda

```{r}
cv.ridge.fit <- cv.glmnet(
  X_train_ridge, 
  y_train_ridge, 
  alpha = 0, 
  family = "binomial", 
  nfolds = 5
) 

lambda_min <- cv.ridge.fit$lambda.min
lambda_min
```

fit using optimal lambda

```{r}
X_test_ridge <- model.matrix(full_predictors_formula, data = test_data)[, -1] 

ridge.probs <- predict(cv.ridge.fit, newx = X_test_ridge, s = lambda_min, type = "response") %>%
  as.vector()

plot(cv.ridge.fit)

ridge.pred <- factor(ifelse(ridge.probs > 0.5, "yes", "no"), levels = c("no", "yes"))
confusionMatrix(ridge.pred, test_data$y, positive = "yes")
```





LASSO
- test across lambda values (and plot) with mean CV test error
- confussion matrix and roc curve

```{r}
X_train_lasso <- X_train_ridge
y_train_lasso <- y_train_ridge
```

k-fold CV for optimal lambda

```{r}
cv.lasso.fit <- cv.glmnet(
  X_train_lasso, 
  y_train_lasso, 
  alpha = 1, 
  family = "binomial", 
  nfolds = 5
) 

lambda_min_lasso <- cv.lasso.fit$lambda.min
lambda_min_lasso


plot(cv.lasso.fit)
title("LASSO CV Error vs. Log(Lambda)", line = 2.5)
abline(v = log(lambda_min_lasso), col = "red", lty = 2) 

```


fit using optimal lambda

```{r}
X_test_lasso <- X_test_ridge # Use the same test design matrix

lasso.probs <- predict(cv.lasso.fit, newx = X_test_lasso, s = lambda_min_lasso, type = "response") %>% 
  as.vector()

lasso.pred <- factor(ifelse(lasso.probs > 0.5, "yes", "no"), levels = c("no", "yes"))
confusionMatrix(lasso.pred, test_data$y, positive = "yes")

```






*plot ROC curves and table confusion matrix for the above models*

```{r}
roc_baseline <- roc(test_data$y, glm.probs.baseline, direction = "<")
roc_ridge <- roc(test_data$y, ridge.probs, direction = "<")


roc_data <- tibble(
  Specificity = c(roc_baseline$specificities, roc_ridge$specificities), # include the raw logit model
  Sensitivity = c(roc_baseline$sensitivities, roc_ridge$sensitivities),
  Model = c(rep(paste0("Baseline (AUC=", round(auc(roc_baseline), 3), ")"), length(roc_baseline$specificities)), 
            rep(paste0("Ridge (AUC=", round(auc(roc_ridge), 3), ")"), length(roc_ridge$specificities)))
)

# Plot the ROC curves 
ggplot(roc_data, aes(x = 1 - Specificity, y = Sensitivity, color = Model)) +
  geom_line(linewidth = 1) +
  geom_abline(lty = 2, color = "gray50") + 
  labs( title = "ROC Curve Comparison: Baseline vs. Ridge Logit",
    x = "False Positive Rate (1 - Specificity)",
    y = "True Positive Rate (Sensitivity)",
    color = "Model") +
  theme_minimal(base_size = 14) 


```



---




## 3.4 K-Nearest Neighbors (KNN)
- Standardize numeric predictors.
- Perform CV to tune **k** .
- Report best k and corresponding CV error.
- Discuss bias-variance behavior.

## 3.5 Generalized Additive Models (GAM)
- Fit GAM with smoothing splines for continuous variables.
- Use CV to choose smoothing parameters.
- Examine partial effect plots for interpretability.

## 3.6 Boosting (Most Flexible Model)
- Fit gradient boosting classifier (e.g., `gbm` or `xgboost`).
- Use cross-validation for:
  - number of trees,
  - interaction depth,
  - learning rate.
- Discuss complexity vs interpretability tradeoff. This model will be our benchmark non-interpretable model to compare interpretable models fitted above.

# 4. Results — Model Assessment & Comparison  

## 4.1 Test Set Performance
For each model:
- Confusion matrix  
  - Accuracy and misclassification rate  
  - Sensitivity / specificity (important for minority class)  
- ROC curves and AUC  

## 4.2 Comparative Evaluation
- Table summarizing all test-set performance metrics.
- Visual comparison of ROC curves.
- Discuss interpretability vs predictive quality.





