---
title: "Statistical Machine Learning Models on A Bank Marketing dataset"
author: "Sandra, Tigist, Mussa Hassen, Waddah"
subtitle: 'Colorado State University, DSCI 445: Statistical Machine Learning, Dr.Andee
  Kaplan'
output:
  ioslides_presentation:
    css: styles.css
---

```{r appendix, include=FALSE}
library(readr)
library(knitr)
library(tidyverse)
library(boot)
library(glmnet)
library(caret)
library(pROC)
library(MASS)
library(ggplot2)

set.seed(445) 

bank_full <- read_delim("data/bank-full_1.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)

bank_full$y <- factor(bank_full$y, levels = c("no", "yes"))


bank_full <- bank_full %>%
  mutate(
    job = as.factor(job),
    marital = as.factor(marital),
    education = as.factor(education),
    default = as.factor(default),
    housing = as.factor(housing),
    loan = as.factor(loan),
    contact = as.factor(contact),
    month = as.factor(month),
    poutcome = as.factor(poutcome)) %>% 
  mutate(previously_contacted = as.factor(ifelse(pdays == -1, "no", "yes")))

train_index <- createDataPartition(bank_full$y, p = 0.8, list = FALSE, times = 1)
train_data <- bank_full[train_index, ]
test_data  <- bank_full[-train_index, ]

cat(paste("Training set size:", nrow(train_data), "\n"))
cat(paste("Test set size:", nrow(test_data), "\n"))


full_formula <- y ~ age + job + marital + education + default + balance + 
  housing + loan + contact + day + month + duration + campaign + 
  pdays + previous + poutcome + previously_contacted

glm.full <- glm(full_formula, data = train_data, family = "binomial")

# Perform backward selection
glm.step <- step(glm.full, direction = "backward", trace = 0)


selected_formula <- formula(glm.step)
selected_formula


glm.baseline <- glm(selected_formula, data = train_data, family = "binomial")
summary(glm.baseline)

glm.tran <- glm(y ~ job + marital + education + housing + loan + contact + 
    day + month + duration + campaign + previous + poutcome + poly(balance, 4), data = train_data, family = "binomial")
summary(glm.tran)


residuals <- residuals(glm.tran, type = "deviance")
train_data$residuals <- residuals

cost <- function(r, pi) {
  mean(r != (pi > 0.5))
}

cv.error.baseline <- cv.glm(
  data = train_data, 
  glmfit = glm.tran, 
  cost = cost, 
  K = 5
)

cv_error_rate <- cv.error.baseline$delta[2]
round(cv_error_rate, 5)

X_full <- model.matrix(y ~ . - 1, data = bank_full)
Y_full <- as.numeric(bank_full$y) - 1 


X_train <- X_full[train_index, ]
Y_train <- Y_full[train_index] 

X_test <- X_full[-train_index, ]
Y_test <- Y_full[-train_index] 

cv.ridge <- cv.glmnet(X_train, Y_train, 
                      alpha = 0, 
                      family = "binomial", 
                      type.measure = "class")
lambda.min.ridge <- cv.ridge$lambda.min


cv.lasso <- cv.glmnet(X_train, Y_train, 
                      alpha = 1, 
                      family = "binomial", 
                      type.measure = "class")
lambda.min.lasso <- cv.lasso$lambda.min


baseline_prob <- predict(glm.tran, newdata = test_data, type = "response")
baseline_pred <- factor(ifelse(baseline_prob > 0.5, "yes", "no"), levels = c("no", "yes"))

ridge_prob <- predict(cv.ridge, s = lambda.min.ridge, newx = X_test, type = "response")
ridge_pred <- factor(ifelse(ridge_prob > 0.5, "yes", "no"), levels = c("no", "yes"))

lasso_prob <- predict(cv.lasso, s = lambda.min.lasso, newx = X_test, type = "response")
lasso_pred <- factor(ifelse(lasso_prob > 0.5, "yes", "no"), levels = c("no", "yes"))

cm_baseline <- confusionMatrix(baseline_pred, test_data$y, positive = "yes")
cm_ridge <- confusionMatrix(ridge_pred, test_data$y, positive = "yes")
cm_lasso <- confusionMatrix(lasso_pred, test_data$y, positive = "yes")

metrics <- data.frame(
  Model = c("Baseline Logit", "Ridge Logit", "LASSO Logit"),
  Accuracy = c(cm_baseline$overall['Accuracy'], cm_ridge$overall['Accuracy'], cm_lasso$overall['Accuracy']),
  Sensitivity = c(cm_baseline$byClass['Sensitivity'], cm_ridge$byClass['Sensitivity'], cm_lasso$byClass['Sensitivity']),
  Specificity = c(cm_baseline$byClass['Specificity'], cm_ridge$byClass['Specificity'], cm_lasso$byClass['Specificity'])
)

roc_baseline <- roc(test_data$y, baseline_prob)
roc_ridge <- roc(test_data$y, as.numeric(ridge_prob))
roc_lasso <- roc(test_data$y, as.numeric(lasso_prob))

roc_data <- data.frame(
  fpr = c(1 - roc_baseline$specificities, 1 - roc_ridge$specificities, 1 - roc_lasso$specificities),
  tpr = c(roc_baseline$sensitivities, roc_ridge$sensitivities, roc_lasso$sensitivities),
  Model = factor(c(rep("Baseline Logit", length(roc_baseline$specificities)),
                   rep("Ridge Logit", length(roc_ridge$specificities)),
                   rep("LASSO Logit", length(roc_lasso$specificities))),
                 levels = c("Baseline Logit", "Ridge Logit", "LASSO Logit"))
)


auc_values <- data.frame(
  Model = c("Baseline Logit", "Ridge Logit", "LASSO Logit"),
  AUC = c(auc(roc_baseline), auc(roc_ridge), auc(roc_lasso))
)


## boosting 

if ("residuals" %in% colnames(train_data)) {
  train_data$residuals <- NULL
}

train_data$y_num <- ifelse(train_data$y == "yes", 1, 0)
test_data$y_num  <- ifelse(test_data$y == "yes", 1, 0)

library(gbm)
set.seed(445)

boost_model <- gbm(
  y_num ~ . - y,         
  data = train_data,
  distribution = "bernoulli",
  n.trees = 1500,       
  shrinkage = 0.01,
  interaction.depth = 3,
  bag.fraction = 0.5,
  verbose = FALSE
)
best_trees <- gbm.perf(boost_model, method = "OOB", plot.it = FALSE)


boost_prob <- predict(boost_model, test_data, n.trees = best_trees, type = "response")
boost_pred <- factor(ifelse(boost_prob > 0.5, "yes", "no"), levels = c("no","yes"))

cm_boost <- confusionMatrix(boost_pred, test_y, positive = "yes")
cm_boost

roc_boost <- roc(test_y, boost_prob)
auc_boost <- auc(roc_boost)

## test 
final_results <- data.frame(
  Model = c("Logistic", "Ridge", "LASSO", "KNN", "GAM", "Boosting"),
  Accuracy = c(
    cm_baseline$overall["Accuracy"],
    cm_ridge$overall["Accuracy"],
    cm_lasso$overall["Accuracy"],
    cm_knn$overall["Accuracy"],
    cm_gam$overall["Accuracy"],
    cm_boost$overall["Accuracy"]
  ),
  Sensitivity = c(
    cm_baseline$byClass["Sensitivity"],
    cm_ridge$byClass["Sensitivity"],
    cm_lasso$byClass["Sensitivity"],
    cm_knn$byClass["Sensitivity"],
    cm_gam$byClass["Sensitivity"],
    cm_boost$byClass["Sensitivity"]
  ),
  AUC = c(
    auc(roc_baseline),
    auc(roc_ridge),
    auc(roc_lasso),
    auc_knn,
    auc_gam,
    auc_boost
  )
)
# combined ROC 
roc_all <- rbind(
  data.frame(fpr = 1 - roc_baseline$specificities, tpr = roc_baseline$sensitivities, Model = "Logistic"),
  data.frame(fpr = 1 - roc_ridge$specificities,    tpr = roc_ridge$sensitivities,    Model = "Ridge"),
  data.frame(fpr = 1 - roc_lasso$specificities,    tpr = roc_lasso$sensitivities,    Model = "LASSO"),
  data.frame(fpr = 1 - roc_knn$specificities,      tpr = roc_knn$sensitivities,      Model = "KNN"),
  data.frame(fpr = 1 - roc_gam$specificities,      tpr = roc_gam$sensitivities,      Model = "GAM"),
  data.frame(fpr = 1 - roc_boost$specificities,    tpr = roc_boost$sensitivities,    Model = "Boosting")
)

```


## Project Motivation & Business Problem (S)
- Telemarketing campaigns are expensive, time-consuming, and have very low success rates. The bank wants to improve efficiency by identifying which clients are most likely to subscribe.
- The dataset has a severe class imbalance: only a small fraction of customers subscribe to the term deposit. This makes prediction difficult and increases the risk of false negatives.
- Improving predictive accuracy would allow the bank to target fewer customers, reduce costs, and increase return on investment (ROI) for future campaigns.


---

```{r, echo=FALSE}
library(ggplot2)

ggplot(bank_full, aes(x = y, fill = y)) +
geom_bar() +
scale_fill_manual(values = c("#2E86C1", "#F39C12")) +
labs(
title = "Term Deposit Subscription (Severe Class Imbalance)",
x = "Subscribed?",
y = "Count"
) +
theme_minimal(base_size = 14)

```


## The Bank Marketing Dataset (Slide 2)
* **Discuss:** The dataset structure: $\approx 45,000$ client observations from a Portuguese banking institution's telemarketing campaigns.
* **Discuss:** Brief overview of the four main predictor categories: Client Demographics, Financial Indicators, Campaign Features, and Past Outcomes. Mention the paper had more features, fro privacy reasons the public data set has less
* **Discuss:** The Target Variable ($y$) is binary: Did the client subscribe to a term deposit (Yes/No)?
* **Discuss:** Emphasize the severity of the class imbalance (e.g., typically $\approx 11\%$ 'Yes' subscriptions). This may yield higher false negatives.
* **Image/Plot Idea:** A small table summarizing the list of the 4 key variable groups (e.g., Demographics, Financial, Campaign, Past). 
* **Image/Plot Idea:** A **Bar Plot** showing the distribution of the target variable ($y$), visually highlighting the imbalance. 



## Problem Statement & Methodology Overview (Slide 3)
* **Discuss:** The project aims for a **Comparative Study**—comparing models based on two criteria: **Predictive Power** and **Interpretability**.
* **Discuss:** The project goal is **Binary Classification**: Building statistical models that are both **predictive** and **interpretable** for management.
* **Discuss:** Established an **80/20 Train/Test Split** and used **5-Fold Cross-Validation (CV)** on the training set for tuning and estimating out-of-sample error.
* **Image/Plot Idea:** An icon representing **Predictive Power** next to an icon representing **Interpretability** (The two core goals).


## Methodology: Data Preparation (Slide 5) (s)

- Converted all categorical variables into factors so models interpret them correctly.
- Addressed "unknown" values by treating them as their own category to avoid losing information.
- Created a new feature previously_contacted to properly encode pdays.
- Split the dataset into 80% training and 20% test using stratified sampling to preserve the class imbalance.
- Ensured numeric predictors were available for later scaling, tuning, and model evaluation.

--- 

## Breakdown of Categorical Variable
```{r, echo=FALSE}
library(ggplot2)

ggplot(bank_full, aes(x = job)) +
geom_bar(fill = "steelblue") +
coord_flip() +
labs(
title = "Distribution of Job Categories",
x = "Job",
y = "Count"
) +
theme_minimal()

```

## LogReg: Feature Selection & Linearity  #MUSSA
* Initial feature selection was performed using **Backward Stepwise Elimination** on the full model to find the most useful set of main effects.
* Our selected formula resutled: 

$y \sim job + marital + education + balance + housing + $

$loan + contact + day + month + duration + $

$campaign + previous+ poutcome$
* **Residual Plots** were used to assess the linearity assumption for numeric predictors (`age`, `balance`). This led to testing a model with non-linear terms.

---

```{r residuals_slide, echo=FALSE, results='asis'}
residuals <- residuals(glm.baseline, type = "deviance")
train_data$residuals <- residuals

ggplot(train_data, aes(x = balance, y = residuals)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess", color = "red", se = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "blue") +
  labs(title = "Deviance Residuals vs. Balance",
       y = "Deviance Residuals") +
  theme_minimal()
```


## LogReg: Models with non-linear components and interactions
* A **quadratic term** $I(\text{balance}^4)$—was added to the formula to reduce the model's bias.
* We then tested for crucial **interaction terms** (between campaign features and demographics). 
* Our model had a cross-validated test error rate 0.09846


## LogReg: Regularization & Tuning   #MUSSA
* We explored Ridge and LASSO regularization as an alternative approach for feature selection.
* **5-Fold Cross-Validation** was used to select the optimal penalty parameter, $\lambda$, which minimized the mean CV error for each regularized model.

```{r, echo=FALSE, results='asis', fig.width=4, fig.height=3}

plot(cv.ridge) 
plot(cv.lasso)
```

## LogReg: Evaluation & Coefficients  #MUSSA
* The final selected LogReg model was evaluated on the 5-fold CV error rate, along with metrics focused on the confusion matrix.
* Models performed relatively similar. We would prefer a model with higher **sensitivity** (few false negatives)

```{r, echo=FALSE, results='asis'}
metrics$AUC <- auc_values[, ncol(auc_values)]
metrics %>% kable()
```

---
# 2. K-Nearest Neighbors (KNN) Model
---

## KNN 1/3: Feature Preparation 
* **Discuss:** **KNN** requires scaling: all numeric predictors were **standardized** (mean=0, SD=1) to prevent features with larger scales from dominating the distance calculation.
* **Discuss:** The non-parametric nature means **no explicit feature selection** via coefficients, but the scaling step is vital for ensuring equal contribution from all dimensions.
* **Image/Plot Idea:** A visual comparison: a scatter plot of two features *unscaled* vs. the same features *scaled/standardized*. 
```{r}
library(dplyr)
library(class)
library(caret)


set.seed(123)
train_index <- sample(nrow(bank_full), 0.8 * nrow(bank_full))
bank_train <- bank_full[train_index, ]
bank_test  <- bank_full[-train_index, ]

numeric_vars <- names(bank_full)[sapply(bank_full, is.numeric)]
numeric_vars <- setdiff(numeric_vars, "y")

train_x <- bank_train[, numeric_vars, drop = FALSE]
test_x  <- bank_test[, numeric_vars, drop = FALSE]

train_x <- bank_train[, numeric_vars]
test_x  <- bank_test[, numeric_vars]

train_x_scaled <- scale(train_x)
test_x_scaled <- scale(
  test_x,
  center = attr(train_x_scaled, "scaled:center"),
  scale  = attr(train_x_scaled, "scaled:scale")
)

train_y <- bank_train$y
test_y  <- bank_test$y

best_k <- 5

par(mfrow = c(1, 2))

plot(
  bank_full$age, bank_full$balance,
  xlab = "Age", ylab = "Balance",
  main = "Unscaled Features"
)

scaled_two <- scale(bank_full[, c("age", "balance")])

plot(
  scaled_two[, "age"], scaled_two[, "balance"],
  xlab = "Scaled Age", ylab = "Scaled Balance",
  main = "Scaled Features (mean = 0, sd = 1)"
)

par(mfrow = c(1, 1))

i <- 1  
test_point <- test_x_scaled[i, , drop = FALSE]

distances <- apply(
  train_x_scaled,
  1,
  function(row) sqrt(sum((row - test_point)^2))
)

nn_idx <- order(distances)[1:best_k]

knn_demo_pred <- knn(
  train = train_x_scaled,
  test  = test_point,
  cl    = train_y,
  k     = best_k
)

cat("Test customer index:", i, "\n")
cat("KNN predicted class:", knn_demo_pred, "\n")
cat("True class:", test_y[i], "\n\n")

cat("Neighbor outcome breakdown:\n")
print(table(train_y[nn_idx]))

cat("\nNearest neighbor details:\n")
print(head(bank_train[nn_idx, c("age", "balance", "duration", "campaign", "y")]))

```
```{r}
library(class)

numeric_vars <- names(bank_full)[sapply(bank_full, is.numeric)]
numeric_vars <- setdiff(numeric_vars, "y")

train_x <- bank_train[, numeric_vars, drop = FALSE]
test_x  <- bank_test[, numeric_vars, drop = FALSE]

train_x_scaled <- scale(train_x)
test_x_scaled <- scale(
  test_x,
  center = attr(train_x_scaled, "scaled:center"),
  scale  = attr(train_x_scaled, "scaled:scale")
)

train_y <- bank_train$y
test_y  <- bank_test$y

knn_pred <- knn(
  train = train_x_scaled,
  test  = test_x_scaled,
  cl    = train_y,
  k     = 5
)

knn_accuracy <- mean(knn_pred == test_y)
knn_error <- 1 - knn_accuracy
cat("KNN Test Accuracy:", round(knn_accuracy * 100, 2), "%\n")
cat("KNN Test Error:", round(knn_error * 100, 2), "%\n")

```
```


## KNN 2/3: Hyperparameter Tuning 
* **Discuss:** The sole hyperparameter, **$k$** (the number of neighbors), determines the model's complexity (i.e., the bias-variance trade-off).
* **Discuss:** **5-Fold Cross-Validation** was performed across a range of $k$ values to find the optimal $k$ that minimized the mean CV error.
* **Image/Plot Idea:** A plot showing **CV Test Error vs. $k$**, illustrating the U-shape curve where the optimal $k$ is selected.
```{r}
library(class)
library(caret)

set.seed(123)

k_values <- 1:25                  
cv_error <- numeric(length(k_values))

folds <- createFolds(train_y, k = 5, returnTrain = FALSE)

for (i in seq_along(k_values)) {
  k <- k_values[i]
  fold_err <- numeric(length(folds))
  
  for (j in seq_along(folds)) {
    test_idx  <- folds[[j]]
    train_idx <- setdiff(seq_along(train_y), test_idx)
    
    fold_train_x <- train_x_scaled[train_idx, , drop = FALSE]
    fold_test_x  <- train_x_scaled[test_idx,  , drop = FALSE]
    fold_train_y <- train_y[train_idx]
    fold_test_y  <- train_y[test_idx]
    
    pred <- knn(
      train = fold_train_x,
      test  = fold_test_x,
      cl    = fold_train_y,
      k     = k
    )
    

    fold_err[j] <- mean(pred != fold_test_y)
  }
  

  cv_error[i] <- mean(fold_err)
}

best_k <- k_values[which.min(cv_error)]
best_k
plot(
  k_values, cv_error,
  type = "b",
  xlab = "Number of neighbors (k)",
  ylab = "5-fold CV Test Error",
  main = "KNN: 5-fold CV Test Error vs k"
)

abline(v = best_k, lty = 2)  

```



## KNN 3/3: Evaluation & Test Performance
* **Discuss:** The final KNN model (with optimal $k$) was applied to the **Test Set**. Evaluate performance using **AUC** and compare it to the LogReg models.
* **Discuss:** KNN generally provides a good intermediate balance—often capturing non-linearity better than LogReg but being less computationally expensive than boosting.
* **Image/Plot Idea:** A brief table comparing the **AUC** and **Recall** of the best LogReg model and the final KNN model.

```{r}
library(pROC)
library(caret)

set.seed(123)

knn_final <- knn(
  train = train_x_scaled,
  test  = test_x_scaled,
  cl    = train_y,
  k     = best_k,
  prob  = TRUE
)

knn_prob_attr <- attr(knn_final, "prob")
knn_prob_yes <- ifelse(knn_final == "yes", knn_prob_attr, 1 - knn_prob_attr)

test_y_fac <- factor(test_y, levels = c("no", "yes"))

knn_roc  <- roc(response = test_y_fac, predictor = knn_prob_yes, levels = c("no", "yes"))
knn_auc  <- as.numeric(auc(knn_roc))

knn_cm      <- confusionMatrix(knn_final, test_y_fac, positive = "yes")
knn_recall  <- knn_cm$byClass["Recall"]



logit_model <- glm(y ~ ., data = bank_train, family = binomial)

logit_prob <- predict(logit_model, newdata = bank_test, type = "response")
logit_pred <- ifelse(logit_prob >= 0.5, "yes", "no")
logit_pred <- factor(logit_pred, levels = c("no", "yes"))

logit_roc <- roc(response = test_y_fac, predictor = logit_prob, levels = c("no", "yes"))
logit_auc <- as.numeric(auc(logit_roc))

logit_cm     <- confusionMatrix(logit_pred, test_y_fac, positive = "yes")
logit_recall <- logit_cm$byClass["Recall"]



results <- data.frame(
  Model  = c("Logistic Regression", paste0("KNN (k = ", best_k, ")")),
  AUC    = c(logit_auc, knn_auc),
  Recall = c(as.numeric(logit_recall), as.numeric(knn_recall))
)

results$AUC    <- round(results$AUC, 3)
results$Recall <- round(results$Recall, 3)

results

```


---
# 3. Boosting Model (High Predictive Benchmark)
---

## Boosting 1/3: Model Formulation
- Gradient Boosting Machines (GBM) build an ensemble of weak learners (typically shallow decision trees).
- Each tree is fit to the errors (residuals) of the previous tree — allowing the model to improve sequentially.
- This iterative correction makes boosting highly flexible and able to capture complex, nonlinear patterns in the data.
- Because the model is effectively a sum of many small trees, it becomes a strong predictor but loses interpretability relative to logistic regression.
- Boosting is expected to be our highest-performing model, especially on imbalanced datasets like this one.




## Boosting 2/3: Hyperparameter Tuning 
- We tuned the number of trees using test-error curves to prevent overfitting.
- Test error decreased steadily and leveled off around M = 1500 trees, which we selected as the optimal value.
- This choice balances model complexity and generalization, giving strong predictive performance without excessive tree growth.


## Boosting 3/3: Evaluation & Importance 

--- 





## Results: Comparative Evaluation (Slide 13). (S)
- We evaluated six models using Accuracy, Sensitivity (Recall), and AUC on the Test Set.
- Boosting achieved the highest overall accuracy (0.905) and the highest sensitivity (0.412), meaning it identified the most true positives.
- Logistic, Ridge, and LASSO models performed very similarly in accuracy (~0.898–0.900) and AUC (~0.907–0.908), but with lower sensitivity.
GAM achieved moderate accuracy and AUC but very low sensitivity.
KNN performed the worst, with the lowest AUC (0.617), indicating weak separation between classes.

--- 

```{r, echo=FALSE}
final_results <- data.frame(
  Model = c("Logistic", "Ridge", "LASSO", "KNN", "GAM", "Boosting"),
  Accuracy = c(
    cm_baseline$overall["Accuracy"],
    cm_ridge$overall["Accuracy"],
    cm_lasso$overall["Accuracy"],
    cm_knn$overall["Accuracy"],
    cm_gam$overall["Accuracy"],
    cm_boost$overall["Accuracy"]
  ),
  Sensitivity = c(
    cm_baseline$byClass["Sensitivity"],
    cm_ridge$byClass["Sensitivity"],
    cm_lasso$byClass["Sensitivity"],
    cm_knn$byClass["Sensitivity"],
    cm_gam$byClass["Sensitivity"],
    cm_boost$byClass["Sensitivity"]
  ),
  AUC = c(
    auc(roc_baseline),
    auc(roc_ridge),
    auc(roc_lasso),
    auc_knn,
    auc_gam,
    auc_boost
  )
)

knitr::kable(final_results, digits = 3)

```
---

```{r, echo=FALSE}
roc_all <- rbind(
  data.frame(fpr = 1 - roc_baseline$specificities, tpr = roc_baseline$sensitivities, Model = "Logistic"),
  data.frame(fpr = 1 - roc_ridge$specificities,    tpr = roc_ridge$sensitivities,    Model = "Ridge"),
  data.frame(fpr = 1 - roc_lasso$specificities,    tpr = roc_lasso$sensitivities,    Model = "LASSO"),
  data.frame(fpr = 1 - roc_knn$specificities,      tpr = roc_knn$sensitivities,      Model = "KNN"),
  data.frame(fpr = 1 - roc_gam$specificities,      tpr = roc_gam$sensitivities,      Model = "GAM"),
  data.frame(fpr = 1 - roc_boost$specificities,    tpr = roc_boost$sensitivities,    Model = "Boosting")
)

ggplot(roc_all, aes(x = fpr, y = tpr, color = Model)) +
  geom_line(size = 1) +
  geom_abline(linetype = "dashed") +
  labs(
    title = "ROC Curve Comparison for All Models",
    x = "False Positive Rate",
    y = "True Positive Rate"
  ) +
  theme_minimal()
```



## The Trade-Off / Business Impact  #MUSSA
* **Discuss:** Reiterate the central trade-off: **Boosting** is the best *predictor*, but the **Regularized Logistic Regression** offers the most *interpretable* and *actionable* insights for the bank.
* **Discuss:** Final actionable advice for the bank: use **LogReg insights** to design campaigns and **Boosting** for a final scoring model if maximum predictive accuracy is needed.
* **Image/Plot Idea:** A visual metaphor (e.g., a scale or seesaw) contrasting "High Performance" vs. "High Interpretability."
* **Discuss:** How the **chosen model** (Balancing performance and interpretability) can lead to **cost savings** and **higher ROI** for future campaigns.
* **Discuss:** Briefly touch upon **regulatory/ethical implications** in banking and the need to justify predictions (further supporting the value of interpretability).
* **Image/Plot Idea:** A currency symbol or a bar chart showing the conceptual increase in ROI due to better targeting.


## Discussion: Limitations & Future Work (Slide 14)  #MUSSA
* **Discuss:** Acknowledge limitations, such as not using **time-series analysis** or fully accounting for the dependency of the 'duration' feature.
* **Discuss:** Suggest future improvements: Exploring advanced techniques like **resampling methods** (SMOTE) to handle imbalance better or trying **Deep Learning** models.


## References (Slide 15)
* **Discuss:** Simply list the required references for the data source and any major methodology papers.
* **Discuss:** Provide contact information for questions.
* Thank you
* **Image/Plot Idea:** The official UCI ML Repository logo and your contact email.



