---
title: "Tree Model"
output: html_document
date: "2024-11-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(ISLR)
library(tidyverse) 
library(tidymodels)
library(knitr)
library(rpart.plot)
library(caret)
library(dplyr)
library(dials)
```


## Packages

- ISLR
- tidyverse
- tidymodels
- knitr
- rpart.plot
- caret
- dplyr
- dials



## Read CSV File

```{r}
drums <- read.csv("sample_features_balanced.csv") |>
  drop_na() |>
  mutate(label = as.factor(label)) 

```



## Initial Tree Model

The drums data set is split into 75% training data, and 25% testing data, stratified by the classification 'label'. A basic tree model is fit using the decision_tree() function with five terminal nodes.  

The tree has four decisions: 
- core frequency < 111
- duration < 1.5
- average activation < 0.0052
- max amplitude < 0.57

```{r}
set.seed(445)

data_split <- initial_split(drums, 0.75, strata = label)
training_drums <- training(data_split)
testing_drums <- testing(data_split)

tree_spec <- decision_tree() |>
  set_engine('rpart') |>
  set_mode("classification")

tree_fit <- tree_spec |>
  fit(label ~. , data = training_drums)

tree_engine <- extract_fit_engine(tree_fit)
rpart.plot(tree_engine, roundint =  FALSE)
```

## Predictions and Accuracy

The initial model has a testing accuracy of 0.86 and correctly classifies:

- 22/26 Hats
- 25/26 Kicks
- 25/26 Rides
- 18/26 Snares

```{r}
set.seed(445)

predictions <- tree_fit |>
  predict(testing_drums) |>
  pull(.pred_class)

metrics <- metric_set(accuracy)

model_performance <- testing_drums %>%
 mutate(predictions = predictions) %>%
 metrics(truth = label, estimate = predictions)

confmatrix <- confusionMatrix(testing_drums$label,predictions)
confmatrix$table


model_performance$.estimate
```

## Tuning Initial Tree Model

We used 10-fold cross-validation to find the best cost_complexity parameter and tree depth. 

- Best CC Paremeter: 0.01
- Best tree depth: 5



```{r}
set.seed(445)

tune_spec <- 
  decision_tree(
    cost_complexity = tune(),
    tree_depth = tune()
  ) %>% 
  set_engine("rpart") %>% 
  set_mode("classification")


tree_grid <- grid_regular(cost_complexity(),
                          tree_depth(),
                          levels = 10)


folds <- vfold_cv(training_drums, v = 10)

tree_wf <- workflow() %>%
  add_model(tune_spec) %>%
  add_formula(label ~ .)

tree_res <- tree_wf %>% 
  tune_grid(resamples = folds,
            grid = tree_grid)



tree_res %>%
  collect_metrics() %>%
  dplyr::filter(.metric != "brier_class") %>%
  dplyr::filter(.metric != "roc_auc") %>%
  mutate(tree_depth = factor(tree_depth)) %>%
  ggplot(aes(cost_complexity, mean, color = tree_depth)) +
  geom_line(linewidth= 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) +
  scale_x_log10(labels = scales::label_number()) +
  scale_color_viridis_d(option = "plasma", begin = .9, end = 0)



best_tree <- tree_res %>%
  select_best(metric="accuracy")

tuned_cc <- best_tree$cost_complexity
tuned_cc
best_tree$tree_depth
```

## Final Tree Model

We refit the tree model using our tuned cost complexity parameter. With the new tuned paramter, we produce the exact same model and predictions as our initial model. 

```{r}
set.seed(445)

tree_spec_tuned <- decision_tree(cost_complexity = tuned_cc) |>
  set_engine('rpart') |>
  set_mode("classification")

tree_fit_tuned <- tree_spec_tuned |>
  fit(label ~. , data = training_drums)


predictions_tuned <- tree_fit_tuned |>
  predict(testing_drums) |>
  pull(.pred_class)

metrics_tuned <- metric_set(accuracy)

model_performance_tuned <- testing_drums %>%
 mutate(predictions = predictions_tuned) %>%
 metrics_tuned(truth = label, estimate = predictions)


model_performance_tuned$.estimate

confmatrix <- confusionMatrix(testing_drums$label,predictions_tuned)
confmatrix$table

tree_engine_tuned <- extract_fit_engine(tree_fit_tuned)
rpart.plot(tree_engine_tuned, roundint =  FALSE)

```

## Discussion

Although our model did not change at all after tuning the cost complexity parameter, we can explain this with our graph of the 10-fold cross validation step. We can see that changing the tree depth does almost nothing to change our accuracy for all trees with depth higher than 2. The cost complexity parameter has no effect on the accuracy until it has a value of 0.01 or higher, where there is a slight increase in accuracy. 













